---
title: "PeerReview_4"
author: "Minxin Cheng"
output: pdf_document
---
0. Load packages
```{r}
#install.packages("tm")
#install.packages("SnowballC")
#install.packages("wordcloud")
#install.packages("e1071")
#install.packages("klaR")

library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
library(klaR)
```

# Problem 1

1. Read in data file
```{r}
sms_raw <- read.csv("da5030.spammsgdataset.csv", stringsAsFactors = FALSE)
```

2. Get the overall information of the data
```{r}
# check the structure of the data
str(sms_raw)
# convert type column to factor
sms_raw$type <- factor(sms_raw$type)
# check the type column and count the number in each type
str(sms_raw$type)
table(sms_raw$type)
```

3. Create corpus
```{r}
# create sms corpus
sms_corpus <- VCorpus(VectorSource(sms_raw$text))
# print the corpus to check
print(sms_corpus)
```

4. Clean the dataset
```{r}
# convert text string to all lower case
sms_corpus_clean <- tm_map(sms_corpus, content_transformer(tolower))
# remove numbers from text
sms_corpus_clean <- tm_map(sms_corpus_clean, removeNumbers)
# remove stop words such as "to", "and", and "or"
sms_corpus_clean <- tm_map(sms_corpus_clean, removeWords, stopwords())
# remove punctuation
sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation)
# only keep the stems of the word
sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
# remove the blank spaces due to the cleaning
sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace)
```

5. Splitting text documents into words, create a sparse matrix
```{r}
sms_dtm <- DocumentTermMatrix(sms_corpus_clean)
```

6. Create training data and testing data
```{r}
# split the data by row number
sms_dtm_train <- sms_dtm[1:4169, ]
sms_dtm_test <- sms_dtm[4170:5559, ]
# get the type information from the original dataset
sms_train_labels <- sms_raw[1:4169, ]$type
sms_test_labels <- sms_raw[4170:5559, ]$type
# check the portion of ham and span in training and testing data
prop.table(table(sms_train_labels))
prop.table(table(sms_test_labels))
```

7. Create word cloud
7.1 Create a word cloud for the entire dataset
```{r}
wordcloud(sms_corpus_clean, min.freq = 50, random.order = FALSE)
```

7.2 Create word cloud for spam data and ham data respectively
```{r}
# separate span and ham data
spam <- subset(sms_raw, type == "spam")
ham <- subset(sms_raw, type == "ham")
# create word cloud for each
wordcloud(spam$text, max.words = 40, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 40, scale = c(2, 0.5))
```
8. Get the most frequent words
```{r}
# only keep the words that appeared at least 5 times
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
# only keep the columns that have the frequent words
sms_dtm_freq_train <- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]
```

9. Convert counts to yes/no for Naive Bayes
```{r}
# create the function
convert_counts <- function(x){
  x <- ifelse(x > 0, "Yes", "No")
}
# apply the function to both training and testing data
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)
```

10. Apply the Naive Bayes
```{r}
sms_classifier <- naiveBayes(sms_train, sms_train_labels)
sms_test_pred <- predict(sms_classifier, sms_test)
```

11. Evaluate the prediction
```{r}
CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c("predicted", "actual"))
```
As the table above, 1390 predictions were made, the accuracy is (1200 + 161) / 1390 = 97.9%

12. Improve the model
```{r}
# set laplace as 1
sms_classifier2 <- naiveBayes(sms_train, sms_train_labels, laplace = 1)
sms_test_pred2 <- predict(sms_classifier2, sms_test)
# evaluate the prediction
CrossTable(sms_test_pred2, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE, dnn = c("predicted", "actual"))
```
As the table above, the accuracy is (1182 + 171) / 1390

# Problem 2
1. Load and check dataset
```{r}
# load data
data(iris)
# get an overview of the data
nrow(iris)
summary(iris)
head(iris)
```

2. split the dataset
```{r}
# get all the row number that can be devided by 5
testidx <- which(1:length(iris[, 1]) %% 5 == 0)
# split the data
iristrain <- iris[-testidx, ]
iristest <- iris[testidx, ]
```

3. Apply Naive Bayes and make the prediction
```{r}
nbmodel <- NaiveBayes(Species~., data = iristrain)
prediction <- predict(nbmodel, iristest[ , -5])
```

4. Evaluate the prediction
```{r}
table(prediction$class, iristest[ , 5])
```
As the table above, the predict accuracy is (10 + 10 + 8) / (10 + 10 + 2 + 8) = 0.93








